---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
activity <- read.csv(unz("activity.zip", "activity.csv"), header=TRUE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r echo=TRUE}
stepsBydate <- aggregate(steps ~ date, activity, sum)
hist(stepsBydate$steps, xlab = "Total Steps per Day", main = "Histogram of Total Steps per Day")
```

```{r echo=TRUE}
meanStepsDate <- as.integer(mean(stepsBydate$steps, na.rm = TRUE))
```

The mean total number of steps taken per day is `r meanStepsDate`.

```{r echo=TRUE}
medianStepsDate <- as.integer(median(stepsBydate$steps, na.rm = TRUE))
```

The median total number of steps taken per day is `r medianStepsDate`.

## What is the average daily activity pattern?

```{r echo=TRUE}
stepsByinterval <- aggregate(steps ~ interval, activity, mean)
plot(stepsByinterval, type = "l", xlab = "Interval (minutes)", ylab = "Average Number of Steps", main = "Average Daily Activity Pattern")
```

```{r echo=TRUE}
maxInterval <- stepsByinterval[stepsByinterval$steps == max(stepsByinterval$steps),]$interval
```
The 5-minute interval, on average across all the days in the dataset, with the maximum number of steps is the `r maxInterval`-minute interval.

## Imputing missing values

```{r echo=TRUE}
nMissing <- sum(is.na(activity[,1]))
```
There are `r nMissing` missing values.

We can fill in the missing values with the mean for that 5-minute interval in a replica of the original dataset.

```{r echo=TRUE}
activity2 <- activity

for (i in 1:nrow(activity2)) {
  if (is.na(activity2[i,1]) == TRUE) {
    activity2[i,1] <- stepsByinterval[which(stepsByinterval$interval == activity2[i,3]),2]
  }
}
```

Below, we have a histogram of the total number of steps taken each day.

```{r echo=TRUE}
stepsBydate2 <- aggregate(steps ~ date, activity2, sum)
hist(stepsBydate2$steps, xlab = "Total Steps per Day", main = "Histogram of Total Steps per Day")
```

```{r echo=TRUE}
meanStepsDate2 <- as.integer(mean(stepsBydate2$steps, na.rm = TRUE))
```

The mean total number of steps taken per day is `r meanStepsDate2`.

```{r echo=TRUE}
medianStepsDate2 <- as.integer(median(stepsBydate2$steps, na.rm = TRUE))
```

The median total number of steps taken per day is `r medianStepsDate2`.

There is a difference in the median total number of steps taken per day from the first part of the assignment, indicating that imputing missing data increases the median.

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
day <- weekdays(activity$date)
activity <- cbind(activity, day)

for (i in 1:nrow(activity)) {
  if (activity[i,4] %in% c("Saturday", "Sunday")) {
    activity[i,4] <- "Weekend"
  }
  else {
    activity[i,4] <- "Weekday"
  }
}

stepsByintervalWeek <- aggregate(steps ~ interval + day, activity, mean)

library(ggplot2)

ggplot(data = stepsByintervalWeek, aes(x = interval, y = steps)) + geom_line() + facet_grid(day ~ .) + ggtitle("Average Daily Activity Pattern") + xlab("Interval (minutes)") + ylab("Average Number of Steps") + theme(plot.title = element_text(hjust = 0.5))
```

From the plot, there are clear differences in average daily activity patterns between weekdays and weekends.